---
title: "Composition"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message = F, warning = F)
options(digits=2)
```

```{r,include=FALSE}
source('libs_init.R')
```

## NMR Metabolomics

Figure 2A in publication

```{r}
df <- bind_rows(
  MEconc_newlong %>% 
    filter(! (str_detect(met,'glycerol') | met == 'ERECTIC' | met=='TSP' | str_detect(met,'PC'))),
  PBconc_newlong %>% filter(! (str_detect(met,'glycerol') | str_detect(met,'EDTA') | met == 'ERECTIC' | met=='TSP' | category=='internal signal' | str_detect(met,'PC') ))
) %>% 
  filter(!is.na(Participant)) %>% 
  group_by(Participant,Cycle,`Cycle day`,`Samples #`) %>% 
  # mutate(y = y / sum(y)) %>% 
  group_by(type,met, SigMa_class,category) %>% 
  dplyr::summarise(mny = sum(y)) %>% 
  group_by(type) %>% 
  dplyr::mutate(mny = mny / sum(mny)) %>% 
  dplyr::mutate(ypos = cumsum(mny) - 0.5*mny, 
                xpos = runif(n(),0.5,1.5)) %>% 
  ungroup() 

df <- df %>% left_join(dflb, by = 'category')
df$lb <- factor(df$lb, levels = dflb$lb[1:9])

```

### Composition 

```{r}
g_composition <- df %>% 
  ggplot(data= ., aes(x = '', y = mny,fill = lb)) + 
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar(theta ="y") +
  theme_void() +
  scale_fill_manual(values=dflb$color)   + 
  facet_wrap(~type) + 
  guides(fill = guide_legend(nrow = 3)) + 
  labs(fill = NULL) + 
  theme(legend.position = 'bottom' )
g_composition


df %>% 
  dplyr::select(type:mny,lb) %>% 
  spread(type,mny) %>% 
  reactable()
```


Sampling time, storage time and batch effect correction of ME-NMR data

```{r}
MEconc_newlong <- MEconc_newlong %>% 
  mutate(Batch = ifelse(is.na(Batch),3,Batch), 
         `Sampling time` = ifelse(is.na(`Sampling time`),'XX',`Sampling time`), 
         `Storage time` = ifelse(is.na(`Storage time`),'XX',`Storage time`)) %>% 
  group_by(met) %>% 
  mutate(logy = residuals(lm(logy~factor(Batch) + factor(`Sampling time`) + factor(`Storage time`)))) %>% 
  ungroup


phyITSold <- phyITS
phyXold <- phyX

MEdf <- MEconc_newlong %>% 
  dplyr::select(`Sample ID`,`Samples #`, met, logy) %>% 
  spread(met,logy) %>% filter(!is.na(`Sample ID`))

MEdfsm <- phyX@sam_data %>% 
  data.frame() %>% 
  dplyr::select(BRK:Medication.during.sampling) %>% 
  left_join(MEdf, by = c('Sample.ID' = 'Sample ID'))  

rownames(MEdfsm) <- sample_names(phyXold)

sample_data(phyITS) <- sample_data(phyX) <- sample_data(MEdfsm)

# save.image(file = '~/Dropbox/Bekzod&Isabella/Yongxin/data/QCcorrected_jan2026.Rdata')

```


## 16s 

### Primers 

Primers for UMI16S:

Primers     - Sequence 

UMI_27Fa    - GTCTCGTGGGCTCGGNNNNNNNNNNNNNNNAGAGTTTGATYMTGGCTYAG
UMI_27Fb    - GTCTCGTGGGCTCGGNNNNNNNNNNNNNNNAGGGTTCGATTCTGGCTCAG
UMI_1540_R  - GTCTCGTGGGCTCGGNNNNNNNNNNNNNNNTACGGYTACCTTGTTACGACT
UMI_338Fa   - GTCTCGTGGGCTCGGNNNNNNNNNNNNNNNACWCCTACGGGWGGCAGCAG
UMI_338Fb   - GTCTCGTGGGCTCGGNNNNNNNNNNNNNNNGACTCCTACGGGAGGCWGCAG
UMI_1391R   - GTCTCGTGGGCTCGGNNNNNNNNNNNNNNNGACGGGCGGTGTGTRCA




Sequencing depth

```{r}
df_ME <- MEconc_newlong %>% filter(!duplicated(`Samples #`))

df_descrip <- phyX@sam_data %>% data.frame()

phyX %>% subset_samples(!is.na(cycle)) %>% sample_sums() %>% summary
```


### Table 

```{r}

TAXlv <- c()
group_vars_list <- list()
for (i in 1:7){
  txlv <- paste('ta',i,sep = '')  
  TAXlv <- c(TAXlv,txlv)
  group_vars_list[[i]] <- TAXlv
}
i <- 0
phyXsummaries <- list()
for (group_vars in group_vars_list) {
  i <- i+1
  txlv <- paste('ta',i,sep = '')  
  phyXtaX <- phyX %>% 
    subset_samples(!is.na(Sample.ID)) %>% 
    transform_sample_counts(function(x) x / sum(x)) %>% 
    taxglom(txlv) %>% 
    psmelt() %>% 
    group_by_at(vars(one_of(group_vars))) %>%
    dplyr::summarise(abu = sum(Abundance), 
                     prevalence = sum(Abundance>0) / n()) %>% 
    ungroup() %>% 
    mutate(abu = abu/ sum(abu), 
           taxlevel = txlv) %>% 
    arrange(-abu)
  phyXsummaries[[i]] <- phyXtaX
}

phyXtaX <- phyX %>% 
  subset_samples(!is.na(Sample.ID)) %>% 
  transform_sample_counts(function(x) x / sum(x)) %>% 
  psmelt() %>% 
  group_by(ta5,ta6,ta7,OTU) %>%
  dplyr::summarise(abu = sum(Abundance), 
                   prevalence = sum(Abundance>0) / n()) %>% 
  ungroup() %>% 
  mutate(abu = abu/ sum(abu), 
         taxlevel = 'ASV') %>% 
  arrange(-abu)

phyXsummaries[[8]] <- phyXtaX

fmrt <- list(
  abu = colDef(format = colFormat(percent = TRUE, digits = 1)),
  prevalence = colDef(format = colFormat(percent = TRUE, digits = 1))
)


phyXsummaries[[1]] %>% head %>% reactable(columns = fmrt)
phyXsummaries[[2]] %>% head %>% reactable(columns = fmrt)
phyXsummaries[[3]] %>% head %>% reactable(columns = fmrt)
phyXsummaries[[4]] %>% head %>% reactable(columns = fmrt)
phyXsummaries[[5]] %>% dplyr::select(-ta1,-ta2) %>%  head %>% reactable(columns = fmrt)
phyXsummaries[[6]] %>% dplyr::select(-ta1,-ta2,-ta3) %>% head %>% reactable(columns = fmrt)
phyXsummaries[[7]] %>% dplyr::select(-ta1,-ta2,-ta3,-ta4) %>% head %>% reactable(columns = fmrt)
phyXsummaries[[8]] %>%  head %>% reactable(columns = fmrt)

```

### Composition 

#### Overall distribution

Figure 2A in publication 

```{r}
b <- phyX %>% 
  subset_samples(!is.na(Sample.ID)) %>% 
  tax_glom('ta5') %>% 
  transform_sample_counts(function(x) x/ sum(x)) %>% 
  psmelt()

bsum <- b %>% 
  group_by(ta2,ta3,ta4,ta5) %>% 
  summarise(mn = mean(Abundance)) %>% 
  ungroup %>% 
  arrange(-mn) %>% 
  mutate(rnk = 1:n(), 
         lb = ifelse(rnk<9,substr(ta5,4,30),'Other'),
         lb = factor(lb)) %>% 
  mutate(mny = mn / sum(mn)) %>% 
  mutate(ypos = cumsum(mn) - 0.5*mn, 
         xpos = runif(n(),0.5,1.5))

bsum$lb <- factor(bsum$lb, levels = bsum$lb[1:9])

g_composition16s <- bsum %>% 
  mutate(omics = 'ME-16s') %>% 
  ggplot(data = ., aes(x = '',y = mn, fill = lb)) + 
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_manual(values=bsum16s$color)   + 
  theme_void() + 
  guides(fill = guide_legend(nrow = 3)) + 
  labs(fill = NULL) + 
  # geom_text(data = bsum[1:8,], aes(x = xpos,y = ypos, label = lb), color = "black", size=3) +
  theme(legend.position = 'bottom' ) + 
  facet_wrap(~omics)
g_composition16s
```

#### Individual distribution

```{r, fig.width=14}
otutb16s <- b %>% left_join(bsum, by = c('ta2','ta3','ta4','ta5')) 
participantsort <- otutb16s %>%
  filter(lb=='Lactobacillaceae') %>% 
  group_by(Participant) %>% 
  dplyr::summarise(ra = mean(Abundance)) %>% 
  arrange(-ra) %>% 
  mutate(rnk = 1:n())

otutb16s %>% 
  mutate(Participant = factor(Participant, levels = participantsort$Participant)) %>% 
  ggplot(data = ., aes(day,Abundance, fill = lb)) + 
  scale_fill_manual(values=bsum16s$color)  + 
  geom_bar(stat = 'identity')  + guides(fill = guide_legend(nrow = 1)) + 
  labs(fill = NULL, x = 'Day in Period') + 
  facet_grid(cycle~Participant , space = 'free', scales = 'free') +
  theme_bw() + 
  theme(legend.position = 'bottom', axis.text.x = element_text(angle = 90))
```


## ITS

### Primers

Primers for ITS sequencing

ITS1 - GTCTCGTGGGCTCGGTCCGTAGGTGAACCTGCGG
ITS4 - GTCTCGTGGGCTCGGTCCTCCGCTTATTGATATGC

### Distribution of zero-reads

```{r}
phyyITS_w <- phyITS %>% subset_samples(!is.na(cycle))
table(phyyITS_w@sam_data$Participant,sample_sums(phyyITS_w)>0)
```

Of the 19 women, 13 have at least a single sample with ITS sequences. 


### Sequencing depth

```{r}
v <- phyITS %>% subset_samples(!is.na(cycle)) %>% sample_sums()
```

For all samples

```{r}
summary(v)
```

For sample with at least one read

```{r}
table(v>0)
summary(v[v>0])
```

### Table 

```{r}

TAXlv <- c()
group_vars_list <- list()
for (i in 1:7){
  txlv <- paste('ta',i,sep = '')  
  TAXlv <- c(TAXlv,txlv)
  group_vars_list[[i]] <- TAXlv
}
i <- 0
phyITSsummaries <- list()
phyITSsummaries_df <- data.frame()
for (group_vars in group_vars_list) {
  i <- i+1
  txlv <- paste('ta',i,sep = '')  
  phyITStaX <- phyITS %>% 
    subset_samples(!is.na(Sample.ID)) %>% 
    transform_sample_counts(function(x) x / sum(x + 1)) %>% 
    taxglom(txlv) %>% 
    psmelt() %>% 
    group_by_at(vars(one_of(group_vars))) %>%
    summarise(abu = sum(Abundance), 
              prevalence = sum(Abundance>0) / n()) %>% 
    ungroup() %>% 
    mutate(abu = abu/ sum(abu), 
           taxlevel = txlv) %>% 
    arrange(-abu)
  phyITSsummaries[[i]] <- phyITStaX
  phyITSsummaries_df %>% bind_rows(phyITSsummaries_df, phyITStaX)
}

phyITStaX <- phyITS %>% 
  subset_samples(!is.na(Sample.ID)) %>% 
  transform_sample_counts(function(x) x / sum(x + 1)) %>% 
  psmelt() %>% 
  group_by(ta4,ta5,ta6,ta7,OTU) %>%
  summarise(abu = sum(Abundance), 
            prevalence = sum(Abundance>0) / n()) %>% 
  ungroup() %>% 
  mutate(abu = abu/ sum(abu), 
         taxlevel = txlv) %>% 
  arrange(-abu)
phyITSsummaries[[8]] <- phyITStaX

phyITSsummaries[[1]] %>% head %>% reactable(columns = fmrt)
phyITSsummaries[[2]] %>% head %>% reactable(columns = fmrt)
phyITSsummaries[[3]] %>% head %>% reactable(columns = fmrt)
phyITSsummaries[[4]] %>% head %>% reactable(columns = fmrt)
phyITSsummaries[[5]] %>% dplyr::select(-ta1,-ta2) %>%  head %>% reactable(columns = fmrt)
phyITSsummaries[[6]] %>% dplyr::select(-ta1,-ta2,-ta3) %>% head %>% reactable(columns = fmrt)
phyITSsummaries[[7]] %>% dplyr::select(-ta1,-ta2,-ta3,-ta4) %>% head %>% reactable(columns = fmrt)
phyITSsummaries[[8]] %>% head %>% reactable(columns = fmrt)

```


### Composition

#### Overall distribution 

Figure 2A in publication 

```{r}
b <- phyITS %>% 
  subset_samples(!is.na(Sample.ID)) %>% 
  tax_glom('ta5') %>% 
  transform_sample_counts(function(x) x/ (sum(x) + 1)) %>% 
  psmelt()

bsum <- b %>% 
  # group_by(ta6) %>% 
  group_by(ta3,ta4,ta5) %>% 
  summarise(mn = sum(Abundance)) %>% 
  ungroup %>% 
  arrange(-mn) %>% 
  mutate(rnk = 1:n(), 
         lb = ifelse(rnk<9,substr(ta5,4,30),'Other'),
         lb = ifelse(is.na(lb),'Other',lb),
         lb = ifelse(ta5 %in% c('F__','__'),paste('Class', substr(ta3,4,30),sep = ': '),lb)) %>% 
  ungroup() %>% 
  mutate(mn = mn / sum(mn))
bsum$lb <- factor(bsum$lb, levels = bsum$lb[1:9])

g_compositionITS <- bsum %>% 
  filter(!is.na(lb)) %>% 
  mutate(omics = 'ME-ITS') %>% 
  ggplot(data = ., aes(x = '',y = mn, fill = lb)) + 
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_manual(values=bsumITS$color) +
  theme_void() + 
  guides(fill = guide_legend(nrow = 3)) + 
  labs(fill = NULL) + 
  # ggrepel::geom_text_repel(data = bsum, aes(x = xpos,y = ypos, label = lb), color = "black", size=3) +
  theme(legend.position = 'bottom' ) + 
  facet_wrap(~omics)

g_compositionITS

```


#### Individual distribution 

```{r, fig.width=14}
otutbITS <- b %>% left_join(bsumITS, by = c('ta3','ta4','ta5')) 
participantsort <- otutbITS %>%
  filter(lb == 'Saccharomycetaceae') %>% 
  group_by(Participant) %>% 
  dplyr::summarise(ra = mean(Abundance)) %>% 
  arrange(-ra) %>% 
  mutate(rnk = 1:n())

otutbITS %>% 
  group_by(Participant,cycle,day,Sample.ID,lb) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  ungroup %>% 
  group_by(Sample.ID) %>% 
  mutate(Abundance = Abundance / sum(Abundance)) %>% 
  mutate(Participant = factor(Participant, levels = participantsort$Participant)) %>% 
  ggplot(data = ., aes(day,Abundance, fill = lb)) + 
  geom_bar(stat = 'identity') + 
  scale_fill_manual(values=bsumITS$color)  + 
  labs(fill = NULL, x = 'Day in Period') + 
  theme_bw() + 
  facet_grid(cycle~Participant) +
  guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = 'bottom', axis.text.x = element_text(angle = 90))

```




## Figure 2 in publication

```{r, fig.width=14}
layout2 <- "
AABC
"

all_comp <- g_composition + g_composition16s + g_compositionITS + 
  plot_layout(design = layout2,guides = "collect") & theme(legend.position = 'bottom',
                                                           legend.direction = 'horizontal')

all_comp
# ggsave('FIG2A_composition.pdf',plot = all_comp, width = 14)
```



## Trajectories over time

### ME

```{r, fig.height=30, fig.width=10}
MEconc_newlong %>% 
  mutate(cycle_day = (as.numeric(Cycle)-1)*5 + `Cycle day`) %>% 
  ggplot(data = ., aes(cycle_day,logy, group = Participant, color = category)) + 
  geom_point() + geom_line() + 
  theme(legend.position = 'bottom') + 
  facet_wrap(~met, ncol = 5, scales = 'free')
```

### 16s




```{r, fig.height=30, fig.width=10}
otutb16s %>% 
  ggplot(data = ., aes(cycle_day,Abundance, group = Participant, color = lb)) + 
  geom_point() + 
  geom_line() + 
  scale_y_log10() + 
  theme(legend.position = 'bottom') + 
  facet_wrap(~OTU + ta5, ncol = 5)
```

### ITS


```{r, fig.height=30, fig.width=10}
otutbITS %>% 
  ggplot(data = ., aes(cycle_day,Abundance, group = Participant, color = lb)) + 
  geom_point() + 
  geom_line() + 
  scale_y_log10() + 
  theme(legend.position = 'bottom') + 
  facet_wrap(~OTU + ta5, ncol = 5)
```

## Effect of day-in-cycle 

### Vulcano plot

Figure 1C in publication 

```{r}
tidyall <- function(m0) {
  r <- bind_rows(
    m0 %>% anova() %>% tidy,
    m0 %>% ranova() %>% tidy,
    m0 %>% tidy)
  return(r)
}

tidyall <- function(fit, data) {
  fit@call$data <- as.data.frame(data)
  
  bind_rows(
    broom::tidy(anova(fit)) %>%
      mutate(test = "fixed"),
    
    broom::tidy(lmerTest::ranova(fit)) %>%
      mutate(test = "random"),
    
    broom.mixed::tidy(fit) %>%
      mutate(test = "coefficient")
  )
}

pc <- 1e-5

tb16s <- phyX %>%
  subset_samples(!is.na(Sample.ID)) %>%
  filter_taxa(function(x) sum(x > 0) > 4, TRUE) %>%
  transform_sample_counts(function(x) x / sum(x)) %>%
  psmelt() %>%
  as.data.frame() %>%
  mutate(lograbu = log(Abundance + pc)) %>% 
  group_by(ta3, ta4, ta5, ta6, ta7, OTU) %>%
  group_modify(~ {
    dat <- as.data.frame(.x)
    
    fit <- lmerTest::lmer(
      lograbu~ factor(day) +
        (1 | Participant) +
        (1 | cycle_part),
      data = dat
    )
    fit@call$data <- dat
    
    # broom::tidy(lmerTest::ranova(fit))
    tidyall(fit, dat)
  }) %>%
  ungroup()



tbITS <- phyITS %>% 
  subset_samples(!is.na(Sample.ID)) %>% 
  filter_taxa(function(x) sum(x>0)>4, TRUE) %>% 
  psmelt() %>% data.frame() %>%
  mutate(lograbu = log(Abundance + pc)) %>% 
  group_by(ta3,ta4,ta5,ta6,ta7, OTU) %>% 
  group_modify(~ {
    dat <- as.data.frame(.x)
    
    fit <- lmerTest::lmer(
      lograbu~ factor(day) +
        (1 | Participant) +
        (1 | cycle_part),
      data = dat
    )
    fit@call$data <- dat
    
    # broom::tidy(lmerTest::ranova(fit))
    tidyall(fit, dat)
  }) %>%
  ungroup()




mME <- MEconc_newlong %>% 
  mutate(cycle_part = paste(Participant, Cycle)) %>% 
  group_by(met,category) %>% 
  group_modify(~ {
    dat <- as.data.frame(.x)
    
    fit <- lmerTest::lmer(
      logy~ factor(`Cycle day`) + (1|`Participant`) + (1|cycle_part),
      data = dat
    )
    fit@call$data <- dat
    tidyall(fit, dat)
  }) %>%
  ungroup()
```


```{r}
dfplot <- bind_rows(mME %>% mutate(omics = 'ME-NMR') %>% 
                      filter(!str_detect(met,'PC')) %>%
                      filter(!str_detect(met,'glycerol')) %>% 
                      filter(!str_detect(met,'TSP')) %>% 
                      filter(!str_detect(met,'ERECTI')) %>% 
                      left_join(dflb, by = 'category'),
                    tbITS %>% mutate(omics = 'ME-ITS', met = OTU) %>% left_join(bsumITS, by = c('ta3','ta4','ta5')),
                    tb16s %>% mutate(omics = 'ME-16s',met = OTU) %>% left_join(bsum16s, by = c('ta3','ta4','ta5'))) %>%  
  filter(str_detect(term,'day'))  %>% 
  filter(!is.na(estimate)) %>% 
  filter(parse_number(term)==2) %>% 
  mutate(omics = factor(omics, levels = c('ME-NMR','ME-16s','ME-ITS'))) 

g_box <- dfplot %>% 
  ggplot(data =., aes(estimate, fill = omics)) + 
  geom_boxplot()+
  facet_wrap(~omics, scales = 'free_x') + 
  scale_fill_manual(values = colorcodes, breaks = colorcodesnames) + 
  theme_bw() +
  theme(legend.position = 'none',
        strip.text = element_blank(),
        strip.background = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank())


g_vulc <- dfplot %>%
  ggplot(data =., aes(estimate,-log10(p.value), label = met, color = lb)) + 
  geom_point() + 
  facet_wrap(~omics, scales = 'free_x') + 
  scale_color_manual(values = bind_rows(dflb, bsum16s,bsumITS)$color, 
                     breaks = bind_rows(dflb, bsum16s,bsumITS)$lb)   + 
  ggrepel::geom_text_repel(data = dfplot[dfplot$p.value<0.01,]) + 
  theme_bw() +
  theme(legend.position = 'none', axis.title.x = element_blank(), axis.text.x = element_blank())

library(patchwork)
layout <- "
A
A
A
A
B
"

g_day2 <- g_vulc + g_box+
  plot_layout(design = layout)
g_day2

```


# Simplex of variance partitioning

Figure 1B in publication

```{r}
tbseq <- bind_rows(tbITS %>% mutate(omics = 'ME-ITS'),
                   tb16s %>% mutate(omics = 'ME-16s')) %>% 
  filter(effect=='ran_pars') %>%
  dplyr::select(-p.value) %>%
  dplyr::mutate(vr = estimate^2) %>% 
  group_by(omics,ta3,ta4,ta5,ta6,ta7,OTU) %>% 
  dplyr::mutate(relvr = vr / sum(vr)) 

tbME <- mME %>% mutate(omics = 'ME-NMR') %>% 
  filter(effect=='ran_pars') %>%
  dplyr::select(-p.value) %>%
  group_by(met,category) %>% 
  dplyr::mutate(vr = estimate^2,
                relvr = vr / sum(vr)) 

brks <- c(0,0.33,0.67,1.00)
gSIMPLX_ME <- tbseq %>% 
  dplyr::select(omics,group,relvr) %>% 
  spread(group,relvr) %>% 
  left_join(bind_rows(bsum16s,bsumITS), by = 'ta5') %>% 
  bind_rows(
    tbME %>% filter(!str_detect(met,'PC')) %>% 
      dplyr::select(omics,group,relvr) %>% 
      spread(group,relvr) %>% left_join(dflb, by = 'category') ) %>% 
  mutate(cycle_part = ifelse(cycle_part==0,0.02,cycle_part), 
         Participant = Participant /(Participant + cycle_part + Residual), 
         Residual = Residual /(Participant + cycle_part + Residual)) %>% 
  ungroup %>% 
  mutate(omics = factor(omics, levels = c('ME-NMR','ME-16s','ME-ITS'))) %>%
  ggtern(data =., aes(cycle_part, Participant, Residual, color = lb)) + 
  geom_point() +
  scale_T_continuous(breaks = brks) +
  scale_L_continuous(breaks = brks) +
  scale_R_continuous(breaks = brks) + 
  scale_color_manual(values = bind_rows(dflb,bsum16s,bsumITS)$color,
                     breaks = bind_rows(dflb,bsum16s,bsumITS)$lb ) +
  xlab('Between Cycles') + 
  ylab('Between Women') + 
  zlab('Within Cycle') + 
  theme(tern.panel.background = element_rect(fill = "white"), 
        tern.panel.grid.major =  element_line(color = "black"),
        tern.axis.arrow.show = TRUE, 
        tern.axis.title.T = element_blank(),
        tern.axis.title.L = element_blank(),
        tern.axis.title.R = element_blank(), 
        legend.position = 'bottom') + 
  facet_wrap(~omics)  

gSIMPLX_ME
```



### Tables

Metabolomics by NMR

```{r}
fmrt <- list(
  # estimate = colDef(format = colFormat(scientific = TRUE, digits = 2)),
  estimate = colDef(
    cell = function(value) {
      if (is.na(value)) return("")
      formatC(value, format = "e", digits = 2)
    }
  ),
  
  std.error = colDef(
    cell = function(value) {
      if (is.na(value)) return("")
      formatC(value, format = "e", digits = 2)
    }
  ),
  
  p.value = colDef(
    cell = function(value) {
      if (is.na(value)) return("")
      formatC(value, format = "e", digits = 2)
    }
  )
)


dfplot %>% 
  ungroup %>% 
  filter(omics=='ME-NMR') %>% 
  filter(!duplicated(met)) %>% 
  arrange(p.value) %>% 
  dplyr::select(met,category,estimate,std.error,p.value) %>% 
  reactable(columns = fmrt)
```

16s microbiome

```{r}
dfplot %>% 
  ungroup %>% 
  filter(omics=='ME-16s') %>% 
  left_join(phyXsummaries[[8]], by = c('met' = 'OTU')) %>% 
  filter(!duplicated(met)) %>% 
  arrange(p.value) %>% 
  dplyr::select(ta4:OTU,estimate,std.error,p.value) %>% 
  reactable(columns = fmrt)
```

ITS Fungome

```{r}
dfplot %>% 
  ungroup %>% 
  filter(omics=='ME-ITS') %>% 
  left_join(phyITSsummaries[[8]], by = c('met' = 'OTU')) %>% 
  filter(!duplicated(met)) %>% 
  arrange(p.value) %>% 
  dplyr::select(ta4.x:OTU,estimate,std.error,p.value) %>% 
  reactable(columns = fmrt)

```

