---
title: "Data"
format:
  html:
    code-fold: false
---

```{r, include=FALSE}
knitr::opts_chunk$set( message = F, warning = F)
options(digits=2)
```

The metabolomics (of menstrual effluent and plasma) as well as 16s and ITS sequencing data, and metadata is available from  https://github.com/mortenarendt/multiomics_menstrual_effluent in the file MEdata.RData. 

* MEconc_new and MEconc_newlong is menstrual effluent metabolomics in wide- and long format
* PBconc_new and PBconc_newlong is plasma metabolomics in wide- and long format
* phyX contains 16s sequencing data 
* phyXITS contains ITS sequencing data
* metadata contrains the sampling time and demographics for the participants
* MEclass contains NMR characterization of the individual intervals

```{r, eval = FALSE, include=TRUE}
url <- paste0(
  "https://raw.githubusercontent.com/",
  "mortenarendt/multiomics_menstrual_effluent/",
  "main/MEdata.RData"
)

tmp <- tempfile(fileext = ".RData")
download.file(url, tmp, mode = "wb")

loaded_objects <- load(tmp)
unlink(tmp)
```

